---
title: "Gov 50 Final Project"
author: "Caroline Gao"
description: "Does a democratic regime’s age strengthen the regime’s electoral democracy?"
output:
  distill::distill_article:
    self_contained: false
---

## Introduction

Does a democratic regime’s age strengthen the regime’s electoral democracy? Through this project, I plan to investigate if and how the age of a country’s current democratic government contributes to the country’s level of electoral democracy. I hypothesize that older democratic regimes (i.e., those founded earlier) will have stronger principles of electoral democracy as measured by the V-Dem electoral democracy index. In other words, the longer a democracy has existed, the higher its electoral democracy index will tend to be. My reasoning behind this hypothesis is that older democracies will have had more time to establish and institutionalize the cultural, historical, and political legitimacy of democratic elections. The longer democratic elections have been taking place within a country, the more embedded they will be in the country’s national identity, value system, and/or political culture, which should theoretically strengthen the principle of electoral democracy within that country. 

A positive correlation between democratic regime age and scores of democratic strength within the data would provide support for my hypothesis. Meanwhile, data showing a negative correlation or zero correlation between these variables would disprove my hypothesis. This question of how a democracy’s age may be related to the strength of the democracy is important as the world reckons with global trends of weakening democracy. If we better understand how the strength of a democracy tends to evolve over time, we can better anticipate what kinds of events or trends might spark future shifts in democratic strength, especially for younger democracies. Moreover, the potential finding that older democracies tend to be stronger would prompt us to investigate what time-related factors may improve the strength and resilience of a given democracy. 




## Data

```{r}
# Load and combine ERT and GDP datasets, subset to democracies observed in the
# most recent ERT update year of 2022, convert GDP into a numerical variable 
# for later regression analysis, sort data alphabetically by country name, and 
# add regime_age variable

library(devtools)
library(tidyverse)

# used this command in console: devtools::install_github("vdeminstitute/ERT")
library(ERT)

gdp_data <- read_csv("~/Gov50/Gov50 Final Project/GDP.csv")

ert <- get_eps() |>
  left_join(gdp_data, by = "country_name") |>
  filter(year == 2022, reg_type == 1) |> 
  mutate(gdp = as.numeric(gdp)) |>
  arrange(country_name) |>
  mutate(regime_age = 2022 - reg_start_year)
```

```{r}
# Create density distribution summarizing the dependent variable, 
# electoral democracy index score, by regime age

score_density <- ert |>
  ggplot(aes(x = v2x_polyarchy)) +
  geom_histogram(aes(y = after_stat(density)), binwidth = 0.05, color = "pink") +
  labs(x = "V-Dem electoral democracy index score", 
       y = "Density",
       title = "Density Histogram of Electoral Democracy Index Score",
       subtitle = "Among Democratic Countries as of 2022",
       caption = "V-Dem Institute, 2022")

score_density
```

For my project, I will primarily analyze V-Dem’s Episodes of Regime Transformation (ERT) data set, which contains 680 unique ERTs from 1900 to 2022. V-Dem defines ERTs as major episodes of regime change toward either democracy or autocracy. I will utilize a cross-sectional observational research design comparing the age and electoral democracy index score of democratic regimes across several countries based on data from 2022 (when the ERT data was most recently updated). The source of this data is the V-Dem Institute, an institute based at the University of Gothenburg that relies on expert coding and granular analysis of country-specific variables to generate globally applicable measures of democracy. 

My explanatory variable of interest is the age of the democratic regime, measured as an integer in years. I will derive this regime age variable from the regime founding year variable (reg_start_year, which is coded as a year between 1900 and 2022) by subtracting the founding year from 2022. V-Dem determined the regime founding year based on their expert analysis of when a given country’s regime transitioned from autocracy to democracy according to V-Dem’s categorization of Regimes of the World. Additionally, I was curious about how GDP may be a confounding variable that increases with time and that may increase the democratic index score of a regime. Thus, I also included a data set from the World Bank on each country’s 2022 GDP in USD to briefly evaluate GDP as an additional potential explanatory variable. 

My outcome variable of interest is the regime’s current score on the V-Dem electoral democracy index (represented by the variable v2x_polyarchy), which is measured as a value in the interval from 0 to 1. This score is determined by the expert coding methodology established by Michael Coppedge et al. via the V-Dem project. To keep my analysis focused on democracies, I will use the regime type variable (reg_type) to filter my data set to only include democracies existing in 2022. As the above density histogram depicts, there is a leftward skew in the distribution of EDIs among democracies in 2022, with the mode of scores being around 0.8.




## Results

# Scatterplot Visualization
```{r}
# Create scatterplot depicting linear regression analysis of
# electoral democracy index score by regime age

age_dem_scatter <- ert |> 
  summarize(country_name, reg_start_year, v2x_polyarchy, regime_age) |>
  ggplot(mapping = aes(x = regime_age, y =  v2x_polyarchy)) +
  geom_point(alpha = 0.8) +
  ylim(0,1) +
  geom_smooth(method = "lm") +
  labs(x = "Regime age (years)", 
       y = "V-Dem electoral democracy index score",
       title = "Electoral Democracy Index Score by Regime Age",
       subtitle = "Among Democratic Countries as of 2022",
       caption = "V-Dem Institute, 2022")

age_dem_scatter
```

This linear regression scatterplot depicts a positive relationship between regime age as an independent variable and electoral democracy index score as a dependent variable. This can be seen in how electoral democracy scores generally increase as regime age increases, resulting in a positively sloped regression line. Of course, it is important to note that this relationship is not perfectly linear or consistent. Interestingly, regimes older than 50 years mostly seem to be clustered either around an age of 75, 100, or 120 years but all tend to have democracy scores around 0.875 (among the highest scores in the data set). This may suggest that the strength of a regime's democracy may plateau and/or become much harder to increase beyond a certain score threshold (which typically is only achieved with greater regime age/maturity).


# Linear Regression
```{r}
# Create linear regression output 

simple_fit <- lm(v2x_polyarchy ~ regime_age, data = ert)

simple_labels <- c(
  "(Intercept)" = "Intercept",
  "regime_age" = "Regime Age (Years)"
)

modelsummary::modelsummary(simple_fit,
                           statistic = c("s.e. = {std.error}",
                                         "p = {p.value}"),
                           coef_map = simple_labels,
                           gof_map = c("nobs", "r.squared", "adj.r.squared"))
```

My linear regression output provides more detail about the nature and strength of the relationship between regime age and electoral democracy index score. The Regime Age coefficient of about 0.003 implies that on average, a 1-year increase in regime age is associated with an average increase of 0.003 points in electoral democracy index score. Meanwhile, the intercept of 0.597 implies that a 0-year-old democracy would be expected to have an average electoral democracy index of 0.597. The r-squared value of about 0.45 suggests that this regression model explains about 0.45 of the observed variability in democracy scores. Finally, the coefficient p-value is very close to zero (<0.001), which suggests that an association this strong or stronger between regime age and electoral democracy index score is statistically significant at a significance level of 0.001 (and perhaps at even lower levels). However, due to the observational, cross-sectional nature of this study, we must first control for potential confounding variables before inferring causality. The following multiple regression analysis will explore the potential confounding role of GDP. 


# Multiple Regression
```{r}
# Create multiple regression output

mult_fit <- lm(v2x_polyarchy ~ regime_age + gdp, data = ert)

mult_labels <- c(
  "(Intercept)" = "Intercept",
  "regime_age" = "Regime Age (Years)",
  "gdp" = "GDP (in USD)"
)

modelsummary::modelsummary(mult_fit,
                           statistic = c("s.e. = {std.error}",
                                         "p = {p.value}"),
                           coef_map = mult_labels,
                           gof_map = c("nobs", "r.squared", "adj.r.squared"))
```

The values in this multiple regression of electoral democracy score based on both regime age and GDP appear surprisingly similar to those in the linear regression. The coefficient of regime age is once again 0.003, which suggests that even when holding GDP constant, a one-year increase in regime age is associated with an average increase in electoral democracy index by 0.003 points. The p-value of such a sizable association is less than 0.001, suggesting that the relationship between regime age and electoral democracy index is statistically significant at a significance level of 0.001 even when controlling for GDP. In terms of interesting outputs besides this main coefficient, the coefficient of GDP being close to zero implies that there is close to no change in the electoral index score when GDP increases by one dollar. Though this near-nonexistent association is not statistically significant, as suggested by the coefficient's p-value of 0.605, it is important to note that countries typically have GDPs in the billions and trillions. Though a single dollar may only be associated with a tiny increase in electoral democracy index, the collective effect of trillions of dollars could lead to consequential trends tying GDP to electoral democracy strength. Thus, even though this multiple regression output suggests GDP doesn't have a statistically significant confounding force on this relationship, I still would not interpret the effects of regime age on electoral democracy index casually. This is because many other potential confounding variables made possible by time, such as more elections or laws securing democracy, may still be at play.


## Conclusion

In summary, my regression analysis provided evidence supporting my hypothesis that older democratic regimes tend to have stronger electoral democracy index scores. Specifically, I found statistically significant positive associations between a democratic regime's age and electoral democracy index, even when controlling for the potential confounding variable of GDP. However, my analysis carries several limitations, one of which was the presence of confounding variables. Due to the cross-sectional nature of my project, I could not control for differences in many country- or region-specific factors affecting the development and strength of democracy. From a more intuitive standpoint, time alone likely cannot strengthen a democracy; rather, the increased number of democratic elections, laws, voters, etc. made possible by the passage of time are more plausible candidates for causing increases in a country's democracy scores. To improve my analysis, I would consider changing the scale of GDP (as used in the multiple regression) to millions or billions of dollars. I would also control for more confounding variables beyond just GDP, including world region, number of democratic laws passed, and population. To better control for country-specific confounders, I would also incorporate analysis of individual countries' democracy scores over time to to see if in-country democracy scores increase with increasing regime age. Finally, I would consider comparing my V-Dem analysis with results from other data sets providing alternative quantitative calculations of democracy. This cross-data comparison would clarify the potential applicability or strength of my conclusions. 
